### 2020-03-24 eca2019proceedings.tex
Import

\documentclass[a4paper,oneside]{book}
\usepackage[utf8x,utf8]{inputenc}
\usepackage{pdfpages}
\usepackage{fancyhdr}
\usepackage{paralist}
\usepackage{tocloft}
\usepackage{imakeidx}
\usepackage[hidelinks]{hyperref}
\usepackage[open,openlevel=1]{bookmark}

\renewcommand{\cftchapleader}{\cftdotfill{\cftdotsep}}

%\usepackage[colorlinks]{hyperref}

\makeindex[name=authors,title=Index of Authors]

\makeatletter
\let\origps@plain\ps@plain
\newcommand\MakePlainPagestyleEmpty{\let\ps@plain\ps@empty}
\newcommand\MakePlainPagestylePlain{\let\ps@plain\origps@plain}
\makeatother

\pagestyle{fancy}
\fancyhf{}% Clear page header/footer
\renewcommand{\headrulewidth}{0pt}% No header rule
\fancyfoot[C]{\thepage}

\newcommand{\insertmydocument}[4]%
           { % Syntax: \insertmydocument{Toc level}{Title}{Subtitle}{File}
             % Requires: tocloft hyperref pdfpages
             \newpage
             \phantomsection
             \cftaddtitleline{toc}{#1}{#2}{\thepage}
             \cftchapterprecistoc{#3}
            \includepdf[pages=-,pagecommand={\pagestyle{fancy}}]{#4}
           }

\makeindex

\begin{document}

\frontmatter
\includepdf[pages=1-,pagecommand={\thispagestyle{empty}}]{layout/cover}
\setcounter{page}{1}
\includepdf[pages=1-,pagecommand={\thispagestyle{empty}}]{layout/firstpages}

\section*{Committees}
	\addcontentsline{toc}{chapter}{Committees}
	\subsection*{General chairs}
	\begin{compactitem}[]
	\item[] Bart Verheij
	\item[] Marco Wiering
	\end{compactitem}

	\subsection*{Contact \& administration}
	\begin{compactitem}[]
	\item[] Elina Sietsema
	\item[] Carlijne de Vries
	\item[] Sarah van Wouwe
	\end{compactitem}

	\subsection*{Local organisation}
	\begin{compactitem}[]
		\item[] Luca Bandelli
		\item[] Abe Brandsma
		\item[] Tomasz Darmetko
		\item[] Mingcheng Ding
		\item[] Ana Dugeniuc
		\item[] Joel During
		\item[] Ameer Islam
		\item[] Siebert Looije
		\item[] Ren\'e Mellema
		\item[] Michaela Mr\`azkov\'a
		\item[] Annet Onnes
		\item[] Benjamin Shafrey
		\item[] Sjaak ten Caat
		\item[] Albert Thie
		\item[] Jelmer van der Linde
		\item[] Luuk van Keeken
		\item[] Paul Veldhuyzen
		\item[] Randy Wind
		\item[] Galiya Yeshmagambetova
	\end{compactitem}

	\subsection*{Program committee}
	\begin{compactitem}[]
		\item[] Stylianos Asteriadis, Maastricht University
		\item[] Reyhan Aydogan, Delft University of Technology
		\item[] Floris Bex, Utrecht University
		\item[] Michael Biehl, University of Groningen
		\item[] Mauro Birattari, Universit\'e Libre de Bruxelles
		\item[] Hendrik Blockeel, Katholieke Universiteit Leuven
		\item[] Hans L. Bodlaender, Utrecht University
		\item[] Sander Bohte, Centrum Wiskunde \& Informatica (CWI)
		\item[] Peter Bosman, Centrum Wiskunde \& Informatica (CWI)
		\item[] Tibor Bosse, Vrije Universiteit Amsterdam
		\item[] Bruno Bouzy, Paris Descartes University
		\item[] Frances Brazier, Delft University of Technology
		\item[] Bert Bredeweg, University of Amsterdam
		\item[] Tristan Cazenave, Universit\'e Paris Dauphine
		\item[] Tom Claassen, Radboud University Nijmegen
		\item[] Walter Daelemans, University of Antwerp
		\item[] Gregoire Danoy, University of Luxembourg
		\item[] Mehdi Dastani, Utrecht University
		\item[] Patrick De Causmaecker, Katholieke Universiteit Leuven
		\item[] Martine De Cock, University of Washington Tacoma
		\item[] Mathijs De Weerdt, Delft University of Technology
		\item[] Beno\^it Depaire, Hasselt University
		\item[] Frank Dignum, Utrecht University
		\item[] Virginia Dignum, Delft University of Technology
		\item[] Kurt Driessens, Maastricht University
		\item[] Madalina Drugan, Technical University of Eindhoven
		\item[] Sjur Dyrkolbotn, Western Norway University of Applied Sciences
		\item[] Jason Farquhar, Radboud University Nijmegen
		\item[] Ad Feelders, Utrecht University
		\item[] Bart Goethals, University of Antwerp
		\item[] Pascal Gribomont, University of Liège
		\item[] Perry Groot, Radboud University Nijmegen
		\item[] Marc Gyssens, Universiteit Hasselt
		\item[] Frank Harmsen, Ernst \& Young Advisory
		\item[] Maaike Harbers, Rotterdam University of Applied Sciences
		\item[] Tom Heskes, Radboud University Nijmegen
		\item[] Koen Hindriks, Delft University of Technology
		\item[] Arjen Hommersom, Open University of the Netherlands
		\item[] Mark Hoogendoorn, Vrije Universiteit Amsterdam
		\item[] Gerda Janssens, Katholieke Universiteit Leuven
		\item[] Maurits Kaptein, Eindhoven University of Technology
		\item[] Uzay Kaymak, Eindhoven University of Technology
		\item[] Walter Kosters, Leiden University
		\item[] Johan Kwisthout, Radboud University Nijmegen
		\item[] Tom Lenaerts, Université Libre de Bruxelles
		\item[] Marco Loog, Delft University of Technology
		\item[] Peter Lucas, Radboud University
		\item[] Elena Marchiori, Radboud University Nijmegen
		\item[] Wannes Meert, Katholieke Universiteit Leuven
		\item[] John-Jules Meyer, Utrecht University
		\item[] Peter Nov\'ak, Delft University of Technology
		\item[] Ann Now\'e, Vrije Universiteit Brussel
		\item[] Aske Plaat, Leiden University
		\item[] Eric Postma, Tilburg University
		\item[] Henry Prakken, University of Utrecht, University of Groningen
		\item[] Jan Ramon, INRIA
		\item[] Silja Renooij, Utrecht University
		\item[] Nico Roos, Maastricht University
		\item[] Stefan Schlobach, Vrije Universiteit Amsterdam
		\item[] Pierre-Yves Schobbens, University of Namur
		\item[] Johannes Scholtes, Maastricht University
		\item[] Lambert Schomaker, University of Groningen
		\item[] Martijn Schut, Universiteit van Amsterdam
		\item[] Evgueni Smirnov, Maastricht University
		\item[] Matthijs T. J. Spaan, Delft University of Technology
		\item[] Gerasimos Spanakis, Maastricht University
		\item[] Jennifer Spenader, University of Groningen
		\item[] Ida Sprinkhuizen-Kuyper, Radboud University Nijmegen
		\item[] Thomas Stützle, Universit\'e Libre de Bruxelles
		\item[] Johan Suykens, Katholieke Universiteit Leuven
		\item[] Niels Taatgen, University of Groningen
		\item[] Annette Ten Teije, Vrije Universiteit Amsterdam
		\item[] Dirk Thierens, Utrecht University
		\item[] Karl Tuyls, University of Liverpool
		\item[] Antal van den Bosch, Radboud University Nijmegen
		\item[] Egon L. van den Broek, Utrecht University
		\item[] Jaap van den Herik, Leiden University
		\item[] Linda van der Gaag, Utrecht University
		\item[] Peter van der Putten, Leiden University, Pegasystems
		\item[] Leon van der Torre, University of Luxembourg
		\item[] Natalie van der Wal, Vrije Universiteit Amsterdam
		\item[] Tim van Erven, Leiden University
		\item[] Marcel van Gerven, Radboud University Nijmegen
		\item[] Frank van Harmelen, Vrije Universiteit Amsterdam
		\item[] Martijn van Otterlo, Vrije Universiteit Amsterdam
		\item[] M. Birna van Riemsdijk, Delft University of Technology
		\item[] Maarten van Someren, University of Amsterdam
		\item[] Marieke van Vugt, University of Groningen
		\item[] Menno van Zaanen, Tilburg University
		\item[] Joaquin Vanschoren, Eindhoven University of Technology
		\item[] Marina Velikova, Radboud University Nijmegen
		\item[] Remco Veltkamp, Utrecht University
		\item[] Joost Vennekens, Katholieke Universiteit Leuven
		\item[] Katja Verbeeck, Technologiecampus Gent
		\item[] Rineke Verbrugge, University of Groningen
		\item[] Michel Verleysen, Universit\'e Catholique de Louvain
		\item[] Sicco Verwer, Delft University of Technology
		\item[] Arnoud Visser, Universiteit van Amsterdam
		\item[] Willem Waegeman, Ghent University
		\item[] Martijn Warnier, Delft University of Technology
		\item[] Gerhard Weiss, Maastricht University
		\item[] Floris Wiesman, Academic Medical Center Amsterdam
		\item[] Jef Wijsen, University of Mons
		\item[] Mark H. M. Winands, Maastricht University
		\item[] Radboud Winkels, Universiteit van Amsterdam
		\item[] Cees Witteveen, Delft University of Technology
		\item[] Marcel Worring, University of Amsterdam
		\item[] Yingqian Zhang, Eindhoven University of Technology
	\end{compactitem}

\include{preface}

\newpage
\pdfbookmark[chapter]{Table of Contents}{toc}
\tableofcontents

\mainmatter
\MakePlainPagestyleEmpty

\include{papers-with-names}

\printindex[authors]

\includepdf[pages=1-,pagecommand={\thispagestyle{empty}}]{layout/back}

\end{document}