### 2020-03-24 papers-with-names.tex

\part{Long papers}

\index[authors]{Aakhus, Mark}
\pdfbookmark[chapter]{Argumentative Design and Polylogue}{paper-long-10}
\insertmydocument{chapter}{Argumentative Design and Polylogue}{Mark Aakhus}{pdfs/long/10longGascón.doc.pdf}

\index[authors]{Bermejo-Luque, Lilian}
\pdfbookmark[chapter]{What is wrong with deductivism?}{paper-long-11}
\insertmydocument{chapter}{What is wrong with deductivism?}{Lilian Bermejo-Luque}{pdfs/long/11LongGoddu.doc.pdf}

\index[authors]{Blair, J. Anthony}
\pdfbookmark[chapter]{Is there an informal logic approach to argument?}{paper-long-12}
\insertmydocument{chapter}{Is there an informal logic approach to argument?}{J. Anthony Blair}{pdfs/long/12longHannkenilljes.docx.pdf}

\index[authors]{Bodlović, Petar}
\pdfbookmark[chapter]{Structural differences between practical and cognitive presumptions}{paper-long-13}
\insertmydocument{chapter}{Structural differences between practical and cognitive presumptions}{Petar Bodlović}{pdfs/long/13longHenning.docx.pdf}

\index[authors]{Cacean, Sebastian}
\pdfbookmark[chapter]{Reliability of Argument Mapping}{paper-long-14}
\insertmydocument{chapter}{Reliability of Argument Mapping}{Sebastian Cacean}{pdfs/long/14longHoffmann.doc.pdf}

\index[authors]{Castro, Diego}
\pdfbookmark[chapter]{Critical discussion for sub-optimal settings}{paper-long-15}
\insertmydocument{chapter}{Critical discussion for sub-optimal settings}{Diego Castro}{pdfs/long/15longHoppmann.doc.pdf}

\index[authors]{Corredor, Cristina}
\pdfbookmark[chapter]{Democratic Legitimacy and Acts of Dissent}{paper-long-16}
\insertmydocument{chapter}{Democratic Legitimacy and Acts of Dissent}{Cristina Corredor}{pdfs/long/16longJackson.doc.pdf}

\index[authors]{Freeman, James}
\pdfbookmark[chapter]{Strength of reasons for moral dissent}{paper-long-18}
\insertmydocument{chapter}{Strength of reasons for moral dissent}{James Freeman}{pdfs/long/18longLewinski.doc.pdf}

\index[authors]{Gascón, José Ángel}
\pdfbookmark[chapter]{Where are dissent and reasons in epistemic justification?}{paper-long-19}
\insertmydocument{chapter}{Where are dissent and reasons in epistemic justification?}{José Ángel Gascón}{pdfs/long/19longLicato.doc.pdf}

\index[authors]{Goddu, G.C.}
\pdfbookmark[chapter]{Justifying Questions?}{paper-long-1}
\insertmydocument{chapter}{Justifying Questions?}{G.C. Goddu}{pdfs/long/1longAakhus.docx.pdf}

\index[authors]{Hannken-Illjes, Kati}
\pdfbookmark[chapter]{Ethnography of Argumentation}{paper-long-20}
\insertmydocument{chapter}{Ethnography of Argumentation}{Kati Hannken-Illjes}{pdfs/long/20longLumer.doc.pdf}

\index[authors]{Henning, Tempest M}
\pdfbookmark[chapter]{“’I said what I said’ - Black women and argumentative politeness norms”}{paper-long-21}
\insertmydocument{chapter}{“’I said what I said’ - Black women and argumentative politeness norms”}{Tempest M Henning}{pdfs/long/21longMosaka.doc.pdf}

\index[authors]{Hoffmann, Michael H.G.}
\pdfbookmark[chapter]{The Argument Assessment Tutor (AAT)}{paper-long-22}
\insertmydocument{chapter}{The Argument Assessment Tutor (AAT)}{Michael H.G. Hoffmann}{pdfs/long/22longReijven.doc.pdf}

\index[authors]{Hoppmann, Michael J.}
\pdfbookmark[chapter]{Grice, Machine Head and the problem of overexpressed premises}{paper-long-23}
\insertmydocument{chapter}{Grice, Machine Head and the problem of overexpressed premises}{Michael J. Hoppmann}{pdfs/long/23longSadek.doc.pdf}

\index[authors]{Zhang, Sally Jackson, Scott Jacobs, Xiaoqi}
\pdfbookmark[chapter]{Standpoints and Commitments as Products of Argumentative Work: Micro/Macro- Analysis of an Infamous Press Conference}{paper-long-24}
\insertmydocument{chapter}{Standpoints and Commitments as Products of Argumentative Work: Micro/Macro- Analysis of an Infamous Press Conference}{Sally Jackson, Scott Jacobs, Xiaoqi Zhang}{pdfs/long/24longSinclair.doc.pdf}

\index[authors]{Lewiński, Marcin}
\pdfbookmark[chapter]{Speech act pluralism in argumentative polylogues}{paper-long-25}
\insertmydocument{chapter}{Speech act pluralism in argumentative polylogues}{Marcin Lewiński}{pdfs/long/25longTindale.docx.pdf}

\index[authors]{Cooper, John Licato, Michael}
\pdfbookmark[chapter]{Evaluating relevance in analogical arguments through warrant-based reasoning}{paper-long-27}
\insertmydocument{chapter}{Evaluating relevance in analogical arguments through warrant-based reasoning}{John Licato, Michael Cooper}{pdfs/long/27longZemplen.doc.pdf}

\index[authors]{Lumer, Christoph}
\pdfbookmark[chapter]{Arguments from expert opinion – An epistemological approach}{paper-long-3}
\insertmydocument{chapter}{Arguments from expert opinion – An epistemological approach}{Christoph Lumer}{pdfs/long/3LongBermejo-Luque.doc.pdf}

\index[authors]{Mosaka, Tshepo Bogosi}
\pdfbookmark[chapter]{All-out attack}{paper-long-4}
\insertmydocument{chapter}{All-out attack}{Tshepo Bogosi Mosaka}{pdfs/long/4longBlair.docx.pdf}

\index[authors]{Reijven, Menno H.}
\pdfbookmark[chapter]{Strategic Maneuvering with Speech Codes: The Rhetorical Use of Cultural Presumptions in Constructing Argumentative Discourse}{paper-long-5}
\insertmydocument{chapter}{Strategic Maneuvering with Speech Codes: The Rhetorical Use of Cultural Presumptions in Constructing Argumentative Discourse}{Menno H. Reijven}{pdfs/long/5longBodlović.doc.pdf}

\index[authors]{Sadek, Karim}
\pdfbookmark[chapter]{Disagreement, public reasoning, and (non-)authoritarian argumentation}{paper-long-6}
\insertmydocument{chapter}{Disagreement, public reasoning, and (non-)authoritarian argumentation}{Karim Sadek}{pdfs/long/6longCacean.docx.pdf}

\index[authors]{Sinclair, Sean}
\pdfbookmark[chapter]{Uncovering Hidden Premises to Reveal the Arguer's Implicit Values: Analysing the Public Debate About Funding Prep}{paper-long-7}
\insertmydocument{chapter}{Uncovering Hidden Premises to Reveal the Arguer's Implicit Values: Analysing the Public Debate About Funding Prep}{Sean Sinclair}{pdfs/long/7longCastro.doc.pdf}

\index[authors]{Tindale, Christopher W.}
\pdfbookmark[chapter]{Strange Fish: Belief and the roots of disagreement}{paper-long-8}
\insertmydocument{chapter}{Strange Fish: Belief and the roots of disagreement}{Christopher W. Tindale}{pdfs/long/8longCorredor.docx.pdf}

\index[authors]{Zemplen, Gabor}
\pdfbookmark[chapter]{Profiling dialogues: Multi-trait mapping of televised argumentative exchanges}{paper-long-9}
\insertmydocument{chapter}{Profiling dialogues: Multi-trait mapping of televised argumentative exchanges}{Gabor Zemplen}{pdfs/long/9longFreeman.doc.pdf}

% \addtocontents{toc}{\protect\newpage}
\part{Regular papers}

\part{Thematic papers}