### 2020-03-24 papers.tex

\part{Long papers}

\insertmydocument{chapter}{Argumentative Design and Polylogue}{Mark Aakhus}{pdfs/long/10longGascón.doc.pdf}
\insertmydocument{chapter}{What is wrong with deductivism?}{Lilian Bermejo-Luque}{pdfs/long/11LongGoddu.doc.pdf}
\insertmydocument{chapter}{Is there an informal logic approach to argument?}{J. Anthony Blair}{pdfs/long/12longHannkenilljes.docx.pdf}
\insertmydocument{chapter}{Structural differences between practical and cognitive presumptions}{Petar Bodlović}{pdfs/long/13longHenning.docx.pdf}
\insertmydocument{chapter}{Reliability of Argument Mapping}{Sebastian Cacean}{pdfs/long/14longHoffmann.doc.pdf}
\insertmydocument{chapter}{Critical discussion for sub-optimal settings}{Diego Castro}{pdfs/long/15longHoppmann.doc.pdf}
\insertmydocument{chapter}{Democratic Legitimacy and Acts of Dissent}{Cristina Corredor}{pdfs/long/16longJackson.doc.pdf}
\insertmydocument{chapter}{Strength of reasons for moral dissent}{James Freeman}{pdfs/long/18longLewinski.doc.pdf}
\insertmydocument{chapter}{Where are dissent and reasons in epistemic justification?}{José Ángel Gascón}{pdfs/long/19longLicato.doc.pdf}
\insertmydocument{chapter}{Justifying Questions?}{G.C. Goddu}{pdfs/long/1longAakhus.docx.pdf}
\insertmydocument{chapter}{Ethnography of Argumentation}{Kati Hannken-Illjes}{pdfs/long/20longLumer.doc.pdf}
\insertmydocument{chapter}{“’I said what I said’ - Black women and argumentative politeness norms”}{Tempest M Henning}{pdfs/long/21longMosaka.doc.pdf}
\insertmydocument{chapter}{The Argument Assessment Tutor (AAT)}{Michael H.G. Hoffmann}{pdfs/long/22longReijven.doc.pdf}
\insertmydocument{chapter}{Grice, Machine Head and the problem of overexpressed premises}{Michael J. Hoppmann}{pdfs/long/23longSadek.doc.pdf}
\insertmydocument{chapter}{Standpoints and Commitments as Products of Argumentative Work: Micro/Macro- Analysis of an Infamous Press Conference}{Sally Jackson, Scott Jacobs, Xiaoqi Zhang}{pdfs/long/24longSinclair.doc.pdf}
\insertmydocument{chapter}{Speech act pluralism in argumentative polylogues}{Marcin Lewiński}{pdfs/long/25longTindale.docx.pdf}
\insertmydocument{chapter}{Evaluating relevance in analogical arguments through warrant-based reasoning}{John Licato, Michael Cooper}{pdfs/long/27longZemplen.doc.pdf}
\insertmydocument{chapter}{Arguments from expert opinion – An epistemological approach}{Christoph Lumer}{pdfs/long/3LongBermejo-Luque.doc.pdf}
\insertmydocument{chapter}{All-out attack}{Tshepo Bogosi Mosaka}{pdfs/long/4longBlair.docx.pdf}
\insertmydocument{chapter}{Strategic Maneuvering with Speech Codes: The Rhetorical Use of Cultural Presumptions in Constructing Argumentative Discourse}{Menno H. Reijven}{pdfs/long/5longBodlović.doc.pdf}
\insertmydocument{chapter}{Disagreement, public reasoning, and (non-)authoritarian argumentation}{Karim Sadek}{pdfs/long/6longCacean.docx.pdf}
\insertmydocument{chapter}{Uncovering Hidden Premises to Reveal the Arguer's Implicit Values: Analysing the Public Debate About Funding Prep}{Sean Sinclair}{pdfs/long/7longCastro.doc.pdf}
\insertmydocument{chapter}{Strange Fish: Belief and the roots of disagreement}{Christopher W. Tindale}{pdfs/long/8longCorredor.docx.pdf}
\insertmydocument{chapter}{Profiling dialogues: Multi-trait mapping of televised argumentative exchanges}{Gabor Zemplen}{pdfs/long/9longFreeman.doc.pdf}

% \addtocontents{toc}{\protect\newpage}
\part{Regular papers}

\part{Thematic papers}