### 2020-03-25 eca2019proceedings.tex
voorwerk

\documentclass[a4paper,oneside]{book}
\usepackage[utf8x,utf8]{inputenc}
\usepackage{pdfpages}
\usepackage{fancyhdr}
\usepackage{paralist}
\usepackage{tocloft}
\usepackage{imakeidx}
\usepackage[hidelinks]{hyperref}
\usepackage[open,openlevel=1]{bookmark}

\renewcommand{\cftchapleader}{\cftdotfill{\cftdotsep}}

%\usepackage[colorlinks]{hyperref}

\makeindex[name=authors,title=Index of Authors]

\makeatletter
\let\origps@plain\ps@plain
\newcommand\MakePlainPagestyleEmpty{\let\ps@plain\ps@empty}
\newcommand\MakePlainPagestylePlain{\let\ps@plain\origps@plain}
\makeatother

\pagestyle{fancy}
\fancyhf{}% Clear page header/footer
\renewcommand{\headrulewidth}{0pt}% No header rule
\fancyfoot[C]{\thepage}

\newcommand{\insertmydocument}[4]%
           { % Syntax: \insertmydocument{Toc level}{Title}{Subtitle}{File}
             % Requires: tocloft hyperref pdfpages
             \newpage
             \phantomsection
             \cftaddtitleline{toc}{#1}{#2}{\thepage}
             \cftchapterprecistoc{#3}
            \includepdf[pages=-,pagecommand={\pagestyle{fancy}}]{#4}
           }

\makeindex

\begin{document}

\frontmatter
\includepdf[pages=1-,pagecommand={\thispagestyle{empty}}]{layout/cover}
\setcounter{page}{1}
\includepdf[pages=1-,pagecommand={\thispagestyle{empty}}]{layout/firstpages}

\include{prep}

\newpage
\pdfbookmark[chapter]{Table of Contents}{toc}
\tableofcontents

\mainmatter
\MakePlainPagestyleEmpty

\include{papers-with-names}

\printindex[authors]

\includepdf[pages=1-,pagecommand={\thispagestyle{empty}}]{layout/back}

\end{document}