### 2020-03-31 papers.tex
Fix most of the names to the point it compiles

% TODO
% [ ] move each commentary directly after each long paper
% [ ] fill in titles of commentaries
% [ ] fill in full author for each commentary
% [ ] Latexify all the weird unicode names (https://en.wikibooks.org/wiki/LaTeX/Special_Characters)

\part{Long papers}

\insertmydocument{chapter}{Argumentative Design and Polylogue}{Mark Aakhus}{pdfs/long/1longAakhus.docx.pdf}
\insertmydocument{chapter}{What is wrong with deductivism?}{Lilian Bermejo-Luque}{pdfs/long/3LongBermejo-Luque.doc.pdf}
\insertmydocument{chapter}{Is there an informal logic approach to argument?}{J. Anthony Blair}{pdfs/long/4longBlair.docx.pdf}
\insertmydocument{chapter}{Structural differences between practical and cognitive presumptions}{Petar Bodlovi\'{c}}{pdfs/long/5longBodlovic.doc.pdf}
\insertmydocument{chapter}{Reliability of Argument Mapping}{Sebastian Cacean}{pdfs/long/6longCacean.docx.pdf}
\insertmydocument{chapter}{Critical discussion for sub-optimal settings}{Diego Castro}{pdfs/long/7longCastro.doc.pdf}
\insertmydocument{chapter}{Democratic Legitimacy and Acts of Dissent}{Cristina Corredor}{pdfs/long/8longCorredor.docx.pdf}
\insertmydocument{chapter}{Strength of reasons for moral dissent}{James Freeman}{pdfs/long/9longFreeman.doc.pdf}
\insertmydocument{chapter}{Where are dissent and reasons in epistemic justification?}{Jos\'{e} \'{A}ngel Gasc\'{o}n}{pdfs/long/10longGascon.doc.pdf}
\insertmydocument{chapter}{Justifying Questions?}{G.C. Goddu}{pdfs/long/11LongGoddu.doc.pdf}
\insertmydocument{chapter}{Ethnography of Argumentation}{Kati Hannken-Illjes}{pdfs/long/12longHannkenilljes.docx.pdf}
\insertmydocument{chapter}{“’I said what I said’ - Black women and argumentative politeness norms”}{Tempest M Henning}{pdfs/long/13longHenning.docx.pdf}
\insertmydocument{chapter}{The Argument Assessment Tutor (AAT)}{Michael H.G. Hoffmann}{pdfs/long/14longHoffmann.doc.pdf}
\insertmydocument{chapter}{Grice, Machine Head and the problem of overexpressed premises}{Michael J. Hoppmann}{pdfs/long/15longHoppmann.doc.pdf}
\insertmydocument{chapter}{Standpoints and Commitments as Products of Argumentative Work: Micro/Macro- Analysis of an Infamous Press Conference}{Sally Jackson, Scott Jacobs, Xiaoqi Zhang}{pdfs/long/16longJackson.doc.pdf}
\insertmydocument{chapter}{Speech act pluralism in argumentative polylogues}{Marcin Lewi\'{n}ski}{pdfs/long/18longLewinski.doc.pdf}
\insertmydocument{chapter}{Evaluating relevance in analogical arguments through warrant-based reasoning}{John Licato, Michael Cooper}{pdfs/long/19longLicato.doc.pdf}
\insertmydocument{chapter}{Arguments from expert opinion – An epistemological approach}{Christoph Lumer}{pdfs/long/20longLumer.doc.pdf}
\insertmydocument{chapter}{All-out attack}{Tshepo Bogosi Mosaka}{pdfs/long/21longMosaka.doc.pdf}
\insertmydocument{chapter}{Strategic Maneuvering with Speech Codes: The Rhetorical Use of Cultural Presumptions in Constructing Argumentative Discourse}{Menno H. Reijven}{pdfs/long/22longReijven.doc.pdf}
\insertmydocument{chapter}{Disagreement, public reasoning, and (non-)authoritarian argumentation}{Karim Sadek}{pdfs/long/23longSadek.doc.pdf}
\insertmydocument{chapter}{Uncovering Hidden Premises to Reveal the Arguer's Implicit Values: Analysing the Public Debate About Funding Prep}{Sean Sinclair}{pdfs/long/24longSinclair.doc.pdf}
\insertmydocument{chapter}{Strange Fish: Belief and the roots of disagreement}{Christopher W. Tindale}{pdfs/long/25longTindale.docx.pdf}
\insertmydocument{chapter}{Profiling dialogues: Multi-trait mapping of televised argumentative exchanges}{Gabor Zemplen}{pdfs/long/27longZemplen.doc.pdf}

% \part{Commentary}

% \insertmydocument{chapter}{TODO-TITLE}{Olmos}{pdfs/commentary/4longBlair-C.doc.pdf}
% \insertmydocument{chapter}{TODO-TITLE}{Freeman}{pdfs/commentary/5longBodlovic-C.doc.pdf}
% \insertmydocument{chapter}{TODO-TITLE}{Pilgram}{pdfs/commentary/6longCacean-C.doc.pdf}
% \insertmydocument{chapter}{TODO-TITLE}{Godden}{pdfs/commentary/7longCastro-C.doc.pdf}
% \insertmydocument{chapter}{TODO-TITLE}{Zarefsky}{pdfs/commentary/8longCorredor-C.doc.pdf}
% \insertmydocument{chapter}{TODO-TITLE}{Casey}{pdfs/commentary/9longFreeman-C.docx.pdf}
% \insertmydocument{chapter}{TODO-TITLE}{Lumer}{pdfs/commentary/10longGascon-C.doc.pdf}
% \insertmydocument{chapter}{TODO-TITLE}{Hitchcock}{pdfs/commentary/11longGoddu-C.docx.pdf}
% \insertmydocument{chapter}{TODO-TITLE}{Hundleby}{pdfs/commentary/13longHenning-C.doc.pdf}
% \insertmydocument{chapter}{TODO-TITLE}{Herman}{pdfs/commentary/15longHoppmann-C.docx.pdf}
% \insertmydocument{chapter}{TODO-TITLE}{Wohlrapp}{pdfs/commentary/16longJackson-C.doc.pdf}
% \insertmydocument{chapter}{TODO-TITLE}{Cruz}{pdfs/commentary/19longLicato-C.docx.pdf}
% \insertmydocument{chapter}{TODO-TITLE}{Innocenti}{pdfs/commentary/22longReijven-C.doc.pdf}
% \insertmydocument{chapter}{TODO-TITLE}{Corredor}{pdfs/commentary/23longSadek-C.doc.pdf}
% \insertmydocument{chapter}{TODO-TITLE}{Jackson}{pdfs/commentary/24longSinclair-C.doc.pdf}
% \insertmydocument{chapter}{TODO-TITLE}{Farine}{pdfs/commentary/25longTindale-C.doc.pdf}
% \insertmydocument{chapter}{TODO-TITLE}{Üzelgün}{pdfs/commentary/27longZemplen-C.doc.pdf}

% \addtocontents{toc}{\protect\newpage}
\part{Regular papers}

\insertmydocument{chapter}{What Optimistic Responses to Deep Disagreement get Right}{Scott F. Aikin}{pdfs/regular/2regularAikin.doc.pdf} % 2
\insertmydocument{chapter}{Is there a crisis of dissent and disagreement in argument theory?}{Mark Alfino}{pdfs/regular/3regularAlfino.docx.pdf} % 3
\insertmydocument{chapter}{Epistemic Tolerance}{Dominik Balg}{pdfs/regular/7regularBalg.doc.pdf} % 7
\insertmydocument{chapter}{Structuring Controversy: The Dialectic of Disagreement}{Sharon Bailin, Mark Battersby}{pdfs/regular/8regularBailin.docx.pdf} % 8
\insertmydocument{chapter}{The role of reasoning and arguing in youth democratic participation in Canada}{Michael D. Baumtrog}{pdfs/regular/9regularBaumtrog.doc.pdf} % 9
\insertmydocument{chapter}{Pictures and Reasoning: visual Arguments and Objections}{Angelina Bobrova}{pdfs/regular/11regularBobrova.docx.pdf} % 11
\insertmydocument{chapter}{Issues leading to argumentative discussions during family mealtime conversations}{Antonio Bova}{pdfs/regular/12regularBova.doc.pdf} % 12
\insertmydocument{chapter}{Value-based argumentation and the Transition to Low Carbon Economy in Turkey and Portugal: Values, Uncertainty and Actions}{Huthaifah Busuulwa}{pdfs/regular/16regularBusuulwa.doc.pdf} % 16
\insertmydocument{chapter}{Dissent: Considering Culture and Personality}{Linda Carozza}{pdfs/regular/17regularCarozza.doc.pdf} % 17
\insertmydocument{chapter}{Heroic Argumentation: On Heroes, Heroism, and Glory in Arguments}{John Casey, Daniel Cohen}{pdfs/regular/18regularCasey.doc.pdf} % 18
\insertmydocument{chapter}{Relationships between narrative and argumentation. In defence of a functional account}{Guillermo Sierra Catal\'{a}n}{pdfs/regular/19regularSierraCatalan.docx.pdf} % 19
\insertmydocument{chapter}{Interactive Discourse Features Supporting Aversive and Existential Acknowledgment In Legislative and Ceremonial Government Apologies}{Martha S. Cheng}{pdfs/regular/20regularCheng.docx.pdf} % 20
\insertmydocument{chapter}{Devil’s Advocates are the Angels of Argumentation}{Katharina Stevens, Daniel Cohen}{pdfs/regular/21regularCohen.docx.pdf} % 21
\insertmydocument{chapter}{The semantic mechanisms underlying disagreement. An argumentative semantics approach}{Ana-Maria Cozma}{pdfs/regular/23regularCozma.docx.pdf} % 23
\insertmydocument{chapter}{Strategic Maneuvering in Implicit an Pseudo-explicit Advertising Discussion}{H\'{e}di Vir\'{a}g Csord\'{a}s}{pdfs/regular/25regularCSORDAS.doc.pdf} % 25
\insertmydocument{chapter}{Reasons for Rational Disagreement from Dialectics The Van Inwagen Cases}{Istvan Danka}{pdfs/regular/26regularDanka.doc.pdf} % 26
\insertmydocument{chapter}{Arguing Brexit on Twitter: a corpus linguistic study}{Natalie Dykes, Philipp Heinrich, Stefan Evert}{pdfs/regular/28regularDykes.docx.pdf} % 28
\insertmydocument{chapter}{Protagoras’ principles, disagreement and the possibility of error}{Michel Dufour}{pdfs/regular/29regularDufour.doc.pdf} % 29
\insertmydocument{chapter}{Changing minds through argumentation: Black Pete as a case study}{Catarina Dutilh Novaes, Emily Sullivan, Thirza Lagewaard, Mark Alfano}{pdfs/regular/30regularDutilhNovaes.doc.pdf} % 30
\insertmydocument{chapter}{The pragma-dialectical view of comparison argumentation}{Frans van Eemeren, Bart Garssen}{pdfs/regular/31regularEemeren.doc.pdf} % 31
\insertmydocument{chapter}{An indexical characterization of disagreement based on possible worlds semantics}{L\'{e}a Farine}{pdfs/regular/35regularFarine.doc.pdf} % 35
\insertmydocument{chapter}{The visual rhetoric of iconic photographs as topoi in editorial cartoons: an argumentative analysis}{Eveline Feteris}{pdfs/regular/36regularFeteris.doc.pdf} % 36
\insertmydocument{chapter}{Conceptual Analysis of an Argumentation using Argumentation Schemes and the Toulmin Model}{Marie Garnier, Patrick Saint-Dizier}{pdfs/regular/37regularGarnier.docx.pdf} % 37
\insertmydocument{chapter}{Representing second-order arguments with Adpositional Argumentation}{Federico Gobbo, Jean H.M. Wagemans}{pdfs/regular/39regularGobbo.doc.pdf} % 39
\insertmydocument{chapter}{Argumentation, dissent, and luck}{Job de Grefte}{pdfs/regular/44regularDeGrefte.doc.pdf} % 44
\insertmydocument{chapter}{Argument Scheme Theory}{Hans V Hansen}{pdfs/regular/45regularHansen.docx.pdf} % 45
\insertmydocument{chapter}{Resolution of deep disagreement: not simply consensus}{Leah Henderson}{pdfs/regular/47regularHenderson.doc.pdf} % 47
\insertmydocument{chapter}{Appeals to Popularity: Roles and Functions of ‘Everyone knows X’}{Thierry Herman}{pdfs/regular/48regularHerman.docx.pdf} % 48
\insertmydocument{chapter}{Towards a Theory of Informal Argument Semantics}{Martin Hinton}{pdfs/regular/49regularHinton.doc.pdf} % 49
\insertmydocument{chapter}{Stereotyping}{David Hitchcock}{pdfs/regular/50regularHitchcock.doc.pdf} % 50
\insertmydocument{chapter}{Changing Norms of Argumentation}{Beth Innocenti}{pdfs/regular/53regularInnocenti.doc.pdf} % 53
\insertmydocument{chapter}{"If You Are A Scientist You Cannot Stop Such A Thing": Scientific Assent and Dissent in the Manhattan Project}{David Erland Isaksen}{pdfs/regular/54regularIsaksen.doc.pdf} % 54
\insertmydocument{chapter}{Formal specifications for dialogue games in multi-party healthcare coaching}{Mathilde Janier, Alison Pease, Mark Snaith, Dominic de Franco}{pdfs/regular/55regularJanier.doc.pdf} % 55
\insertmydocument{chapter}{How courts should respond to the stories defendants tell: A Bayesian account of a Dutch ruling}{Hylke Jellema}{pdfs/regular/57regularJellema.doc.pdf} % 57
\insertmydocument{chapter}{The spaces and places of argumentation: Developing a spatial model of argumentation}{Steven Johnson}{pdfs/regular/58regularJohnson.docx.pdf} % 58
\insertmydocument{chapter}{Aggressiveness as a qualifier of dissent in interpersonal arguing}{Iryna Khomenko}{pdfs/regular/59regularKhomenko.docx.pdf} % 59
\insertmydocument{chapter}{Auditory Arguments – Importance of Sound in an Argumentative Discourse}{Gabrijela Ki\v{s}i\v{c}ek}{pdfs/regular/60regularKisicek.docx.pdf} % 60
\insertmydocument{chapter}{Be reasonable! Ways to react to cases of presumed unreason}{Erik C. W. Krabbe, Jan Albert van Laar}{pdfs/regular/65regularKrabbe.doc.pdf} % 65
\insertmydocument{chapter}{Epistemic Injustice and Deep Disagreement}{Thirza Lagewaard}{pdfs/regular/69regularLagewaard.doc.pdf} % 69
\insertmydocument{chapter}{Argumentative Patterns of Right-Wing Populism: The Example of the “Alternative for Germany”}{David Lanius}{pdfs/regular/70regularLanius.doc.pdf} % 70
\insertmydocument{chapter}{A decision tree for annotating argumentation scheme corpora}{John Lawrence, Jacky Visser, Douglas Walton, Chris Reed}{pdfs/regular/71regularLawrence.doc.pdf} % 71
\insertmydocument{chapter}{Philosophical Foundation of Reasonableness in Mencius’s Argumentative Discourse: Based on the use of dissociation}{Yan Linqiong, Xiong Minghui}{pdfs/regular/72regularYanXiong.docx.pdf} % 72
\insertmydocument{chapter}{Strategic manoeuvring with the construction “We know that X” in parliamentary debates}{Matthijs Looij, Bertijn van der Steenhoven}{pdfs/regular/74regularLooij.docx.pdf} % 74
\insertmydocument{chapter}{Argument, justification and inquiry: a pragmatist approach}{Federico Ezequiel L\'{o}pez}{pdfs/regular/75regularLopez.doc.pdf} % 75
\insertmydocument{chapter}{Dissent as a Voice of Support: The Influence of Argument Types in Providing Support to Those Who have Experienced Racial Discrimination}{Uttara Manohar, Susan L. Kline}{pdfs/regular/78regularManohar.doc.pdf} % 78
\insertmydocument{chapter}{Holism of reasons and its consequences for argumentation theory}{Hubert Marraud}{pdfs/regular/80regularMarraud.doc.pdf} % 80
\insertmydocument{chapter}{“Why did not your correspondent make an honest inquiry before so writing?”: The text structure and discourse of disagreement in Irish letters to the editor}{Davide Mazzi}{pdfs/regular/81regularMazzi.doc.pdf} % 81
\insertmydocument{chapter}{An Argumentative Approach to the Justification of Abduction}{Paula Olmos}{pdfs/regular/86regularOlmos.doc.pdf} % 86
\insertmydocument{chapter}{When You Can’t Trust What You Don’t Know: organ and tissue donation, public trust, the voice of dissent and informed consent}{Harmony Peach}{pdfs/regular/91regularPeach.doc.pdf} % 91
\insertmydocument{chapter}{“The worst ever conceived by a man of genius. ” Hume’s probability argument in A Treatise}{Jeanne Peijnenburg, David Atkinson}{pdfs/regular/92regularPeijnenburg.docx.pdf} % 92
\insertmydocument{chapter}{Character, Dog Whistles, and the Limits of Charity}{Kathryn Phillips}{pdfs/regular/94regularPhillips.doc.pdf} % 94
\insertmydocument{chapter}{The rhetorical potential of metaphor: An experimental study on the effect of metaphors in argumentation}{Roosmaryn Pilgram}{pdfs/regular/97regularPilgram.docx.pdf} % 97
\insertmydocument{chapter}{Separate opinions as an argumentative activity type}{H. Jos\'{e} Plug}{pdfs/regular/98regularPlug.doc.pdf} % 98
\insertmydocument{chapter}{Can Literary Fiction be Suppositional Reasoning?}{Gilbert Plumer}{pdfs/regular/99regularPlumer.doc.pdf} % 99
\insertmydocument{chapter}{Conspiracy Theories and Reasonable Disagreement}{Juho Ritola}{pdfs/regular/106regularRitola.doc.pdf} % 106
\insertmydocument{chapter}{Inference to the Best Metaphor}{Cristi\'{a}n Santib\'{a}\~{n}ez}{pdfs/regular/107regularSantibanez.docx.pdf} % 107
\insertmydocument{chapter}{Beyond Randomized Clinical Trials: Emerging Innovations in Reasoning About Health}{Jodi Schneider, Sally Jackson}{pdfs/regular/109regularSchneider.docx.pdf} % 109
\insertmydocument{chapter}{Multimodal argumentation and dissent – a perspective of multimodal critical discourse analysis}{Andrea Sabine Sedlaczek}{pdfs/regular/111regularSedlaczek.doc.pdf} % 111
\insertmydocument{chapter}{Analogical Argumentation in Philosophical Thought Experiments}{Stefan Sleeuw}{pdfs/regular/112regularSleeuw.docx.pdf} % 112
\insertmydocument{chapter}{Detecting and handling disagreement in multi-party health coaching}{Mark Snaith, Alison Pease, Dominic De Franco}{pdfs/regular/113regularSnaith.doc.pdf} % 113
\insertmydocument{chapter}{Explaining away and the propensity interpretation of probability: the case of unequal priors}{Alice Liefgreen, Marko Te\v{s}i\'{c}}{pdfs/regular/116regularLiefgreen.docx.pdf} % 116
\insertmydocument{chapter}{Responsible Agent Deliberation}{Douglas Walton, Alice Toniolo}{pdfs/regular/118regularToniolo.doc.pdf} % 118
\insertmydocument{chapter}{Disagreement in Aristotle’s Topics and its Latin Medieval reception}{Gustavo Fern\'{a}ndez Walker, Ana Mar\'{i}a Mora-M\'{a}rquez}{pdfs/regular/122regularWalker.docx.pdf} % 122
\insertmydocument{chapter}{Conductive Arguments. Towards a solution of the riddles}{Harald R. Wohlrapp}{pdfs/regular/125regularWohlrapp.doc.pdf} % 125
\insertmydocument{chapter}{A ludological perspective on the shape of argument: Collaborative assent to dissenting opposition}{Michael A. Yong-Set}{pdfs/regular/126regularYong-Set.docx.pdf} % 126
\insertmydocument{chapter}{Towards the justification of warrant in a legal hard case: How to justify a legal rule by the backing}{Shiyang Yu, Xi Chen}{pdfs/regular/127regularYu.docx.pdf} % 127
\insertmydocument{chapter}{How Should We Classify Argument Schemes?}{David Zarefsky}{pdfs/regular/129regularZarefsky.doc.pdf} % 129

\part{Thematic papers}

\insertmydocument{chapter}{Visual and Auditory Schemes of Argument}{Leo Groarke}{pdfs/thematic/1regularthematicGroarke.docx.pdf}
\insertmydocument{chapter}{The prospects for multimodal schemes of argument: Assessing Adbusters’ spoofing strategies}{Assimakis Tseronis}{pdfs/thematic/3regularThematicTseronis.doc.pdf}
\insertmydocument{chapter}{(Mis)framing photographs as an obstacle to fair dissent}{Marta Zampa, Chiara Pollaroli}{pdfs/thematic/4regularThematicZampa.docx.pdf}
\insertmydocument{chapter}{Dissent, Disadvantage, Testimony and the Ideological ‘Truth’ of Presumptions}{Harmony Peach}{pdfs/thematic/5regularThematicPeach.doc.pdf}
\insertmydocument{chapter}{With All Due Respect: Controversial Beliefs and the Limits of Responsible Argumentation}{Tracy Bowell}{pdfs/thematic/7regularThematicBowell.doc.pdf}
\insertmydocument{chapter}{A Quantitative Corpus-Based Study of Evidentiality and Disagreement in Earnings Conference Calls}{Andrea Rocci, Carlo Raimondo, Daniele Puccinelli}{pdfs/thematic/9regularThematicRocci.docx.pdf}
\insertmydocument{chapter}{Five approaches to argument strength: probabilistic, dialectical, structural, empirical, and computational}{Frank Zenker, David Godden, Marcin Selinger, Simon Wells, Kamila Debowska-Kozlowska}{pdfs/thematic/13regularThematicZenkeretal.docx.pdf}
\insertmydocument{chapter}{Responsible Agent Deliberation}{Douglas Walton, Alice Toniolo}{pdfs/thematic/18regularThematicWalton.doc.pdf}
\insertmydocument{chapter}{From dialogic to argument-based teaching: Significant teacher moves in framing classroom-based argumentation dialogues}{Chrysi Rapanta, Dilar Cascalheira}{pdfs/thematic/20regularThematicRapanta.doc.pdf}