### 2020-04-19 papers-vol-2.tex

\part{Regular papers}

\insertmydocument{chapter}{What Optimistic Responses to Deep Disagreement get Right}{Scott F. Aikin}{pdfs/regular/2regularAikin.doc.pdf} % 2
\insertmydocument{chapter}{Is There a Crisis of Dissent and Disagreement in Argument Theory?}{Mark Alfino}{pdfs/regular/3regularAlfino.docx.pdf} % 3
\insertmydocument{chapter}{Epistemic Tolerance}{Dominik Balg}{pdfs/regular/7regularBalg.doc.pdf} % 7
\insertmydocument{chapter}{Structuring Controversy: The Dialectic of Disagreement}{Sharon Bailin, Mark Battersby}{pdfs/regular/8regularBailin.docx.pdf} % 8
\insertmydocument{chapter}{The Role of Reasoning and Arguing in Youth Democratic Participation in Canada}{Michael D. Baumtrog}{pdfs/regular/9regularBaumtrog.doc.pdf} % 9
\insertmydocument{chapter}{Pictures and Reasoning: Visual Arguments and Objections}{Angelina Bobrova}{pdfs/regular/11regularBobrova.docx.pdf} % 11
\insertmydocument{chapter}{Issues Leading to Argumentative Discussions During Family Mealtime Conversations}{Antonio Bova}{pdfs/regular/12regularBova.doc.pdf} % 12
\insertmydocument{chapter}{Value-based Argumentation and the Transition to Low Carbon Economy in Turkey and Portugal: Values, Uncertainty and Actions}{Huthaifah Busuulwa}{pdfs/regular/16regularBusuulwa.doc.pdf} % 16
\insertmydocument{chapter}{Dissent: Considering Culture and Personality}{Linda Carozza}{pdfs/regular/17regularCarozza.doc.pdf} % 17
\insertmydocument{chapter}{Heroic Argumentation: On Heroes, Heroism, and Glory in Arguments}{John Casey, Daniel Cohen}{pdfs/regular/18regularCasey.doc.pdf} % 18
\insertmydocument{chapter}{Relationships between Narrative and Argumentation. In Defence of a Functional Account}{Guillermo Sierra Catal\'{a}n}{pdfs/regular/19regularSierraCatalan.docx.pdf} % 19
\insertmydocument{chapter}{Interactive Discourse Features Supporting Aversive and Existential Acknowledgment In Legislative and Ceremonial Government Apologies}{Martha S. Cheng}{pdfs/regular/20regularCheng.docx.pdf} % 20
\insertmydocument{chapter}{Devil’s Advocates are the Angels of Argumentation}{Katharina Stevens, Daniel Cohen}{pdfs/regular/21regularCohen.docx.pdf} % 21
\insertmydocument{chapter}{The Semantic Mechanisms Underlying Disagreement. An Argumentative Semantics Approach}{Ana-Maria Cozma}{pdfs/regular/23regularCozma.docx.pdf} % 23
\insertmydocument{chapter}{Strategic Maneuvering in Implicit an Pseudo-explicit Advertising Discussion}{H\'{e}di Vir\'{a}g Csord\'{a}s}{pdfs/regular/25regularCSORDAS.doc.pdf} % 25
\insertmydocument{chapter}{Reasons for Rational Disagreement from Dialectics The Van Inwagen Cases}{Istvan Danka}{pdfs/regular/26regularDanka.doc.pdf} % 26
\insertmydocument{chapter}{Arguing Brexit on Twitter: A Corpus Linguistic Study}{Natalie Dykes, Philipp Heinrich, Stefan Evert}{pdfs/regular/28regularDykes.docx.pdf} % 28
\insertmydocument{chapter}{Protagoras’ Principles, Disagreement and the Possibility of Error}{Michel Dufour}{pdfs/regular/29regularDufour.doc.pdf} % 29
\insertmydocument{chapter}{Changing Minds through Argumentation: Black Pete as a Case Study}{Catarina Dutilh Novaes, Emily Sullivan, Thirza Lagewaard, Mark Alfano}{pdfs/regular/30regularDutilhNovaes.doc.pdf} % 30
\insertmydocument{chapter}{The Pragma-dialectical View of Comparison Argumentation}{Frans van Eemeren, Bart Garssen}{pdfs/regular/31regularEemeren.doc.pdf} % 31
\insertmydocument{chapter}{An Indexical Characterization of Disagreement Based on Possible Worlds Semantics}{L\'{e}a Farine}{pdfs/regular/35regularFarine.doc.pdf} % 35
\insertmydocument{chapter}{The Visual Rhetoric of Iconic Photographs as Topoi in Editorial Cartoons: an Argumentative Analysis}{Eveline Feteris}{pdfs/regular/36regularFeteris.doc.pdf} % 36
\insertmydocument{chapter}{Conceptual Analysis of an Argumentation using Argumentation Schemes and the Toulmin Model}{Marie Garnier, Patrick Saint-Dizier}{pdfs/regular/37regularGarnier.docx.pdf} % 37
\insertmydocument{chapter}{Representing Second-order Arguments with Adpositional Argumentation}{Federico Gobbo, Jean H.M. Wagemans}{pdfs/regular/39regularGobbo.doc.pdf} % 39
\insertmydocument{chapter}{Argumentation, Dissent, and Luck}{Job de Grefte}{pdfs/regular/44regularDeGrefte.doc.pdf} % 44
\insertmydocument{chapter}{Argument Scheme Theory}{Hans V Hansen}{pdfs/regular/45regularHansen.docx.pdf} % 45
\insertmydocument{chapter}{Resolution of Deep Disagreement: Not Simply Consensus}{Leah Henderson}{pdfs/regular/47regularHenderson.docx.pdf} % 47
\insertmydocument{chapter}{Appeals to Popularity: Roles and Functions of ‘Everyone knows X’}{Thierry Herman}{pdfs/regular/48regularHerman.docx.pdf} % 48
\insertmydocument{chapter}{Towards a Theory of Informal Argument Semantics}{Martin Hinton}{pdfs/regular/49regularHinton.doc.pdf} % 49
\insertmydocument{chapter}{Stereotyping}{David Hitchcock}{pdfs/regular/50regularHitchcock.doc.pdf} % 50
\insertmydocument{chapter}{Changing Norms of Argumentation}{Beth Innocenti}{pdfs/regular/53regularInnocenti.doc.pdf} % 53
\insertmydocument{chapter}{"If You Are A Scientist You Cannot Stop Such A Thing": Scientific Assent and Dissent in the Manhattan Project}{David Erland Isaksen}{pdfs/regular/54regularIsaksen.doc.pdf} % 54
\insertmydocument{chapter}{Formal Specifications for Dialogue Games in Multi-party Healthcare Coaching}{Mathilde Janier, Alison Pease, Mark Snaith, Dominic de Franco}{pdfs/regular/55regularJanier.docx.pdf} % 55
\insertmydocument{chapter}{How Courts Should Respond to the Stories Defendants Tell: A Bayesian Account of a Dutch Ruling}{Hylke Jellema}{pdfs/regular/57regularJellema.doc.pdf} % 57
\insertmydocument{chapter}{The Spaces and Places of Argumentation: Developing a Spatial Model of Argumentation}{Steven Johnson}{pdfs/regular/58regularJohnson.docx.pdf} % 58
\insertmydocument{chapter}{Aggressiveness as a Qualifier of Dissent in Interpersonal Arguing}{Iryna Khomenko}{pdfs/regular/59regularKhomenko.docx.pdf} % 59
\insertmydocument{chapter}{Auditory Arguments – Importance of Sound in an Argumentative Discourse}{Gabrijela Ki\v{s}i\v{c}ek}{pdfs/regular/60regularKisicek.docx.pdf} % 60
\insertmydocument{chapter}{Be reasonable! Ways to React to Cases of Presumed Unreason}{Erik C. W. Krabbe, Jan Albert van Laar}{pdfs/regular/65regularKrabbe.doc.pdf} % 65
\insertmydocument{chapter}{Epistemic Injustice and Deep Disagreement}{Thirza Lagewaard}{pdfs/regular/69regularLagewaard.doc.pdf} % 69
\insertmydocument{chapter}{Argumentative Patterns of Right-Wing Populism: The Example of the “Alternative for Germany”}{David Lanius}{pdfs/regular/70regularLanius.doc.pdf} % 70
\insertmydocument{chapter}{A Decision Tree for Annotating Argumentation Scheme Corpora}{John Lawrence, Jacky Visser, Douglas Walton, Chris Reed}{pdfs/regular/71regularLawrence.doc.pdf} % 71
\insertmydocument{chapter}{Philosophical Foundation of Reasonableness in Mencius’s Argumentative Discourse: Based on the use of dissociation}{Yan Linqiong, Xiong Minghui}{pdfs/regular/72regularYanXiong.docx.pdf} % 72
\insertmydocument{chapter}{Strategic Manoeuvring with the Construction “We know that X” in Parliamentary Debates}{Matthijs Looij, Bertijn van der Steenhoven}{pdfs/regular/74regularLooij.docx.pdf} % 74
\insertmydocument{chapter}{Argument, Justification and Inquiry: a Pragmatist Approach}{Federico Ezequiel L\'{o}pez}{pdfs/regular/75regularLopez.doc.pdf} % 75
\insertmydocument{chapter}{Dissent as a Voice of Support: The Influence of Argument Types in Providing Support to Those Who have Experienced Racial Discrimination}{Uttara Manohar, Susan L. Kline}{pdfs/regular/78regularManohar.doc.pdf} % 78
\insertmydocument{chapter}{Holism of Reasons and its Consequences for Argumentation Theory}{Hubert Marraud}{pdfs/regular/80regularMarraud.doc.pdf} % 80
\insertmydocument{chapter}{“Why did not your correspondent make an honest inquiry before so writing?”: The Text Structure and Discourse of Disagreement in Irish letters to the Editor}{Davide Mazzi}{pdfs/regular/81regularMazzi.doc.pdf} % 81
\insertmydocument{chapter}{An Argumentative Approach to the Justification of Abduction}{Paula Olmos}{pdfs/regular/86regularOlmos.doc.pdf} % 86
\insertmydocument{chapter}{When You Can’t Trust What You Don’t Know: Organ and Tissue Donation, Public Trust, the Voice of Dissent and Informed Consent}{Harmony Peach}{pdfs/regular/91regularPeach.doc.pdf} % 91
\insertmydocument{chapter}{“The worst ever conceived by a man of genius. ” Hume’s Probability Argument in A Treatise}{Jeanne Peijnenburg, David Atkinson}{pdfs/regular/92regularPeijnenburg.docx.pdf} % 92
\insertmydocument{chapter}{Character, Dog Whistles, and the Limits of Charity}{Kathryn Phillips}{pdfs/regular/94regularPhillips.doc.pdf} % 94
\insertmydocument{chapter}{The Rhetorical Potential of Metaphor: An Experimental Study on the Effect of Metaphors in Argumentation}{Roosmaryn Pilgram}{pdfs/regular/97regularPilgram.docx.pdf} % 97
\insertmydocument{chapter}{Separate Opinions as an Argumentative Activity Type}{H. Jos\'{e} Plug}{pdfs/regular/98regularPlug.doc.pdf} % 98
\insertmydocument{chapter}{Can Literary Fiction be Suppositional Reasoning?}{Gilbert Plumer}{pdfs/regular/99regularPlumer.doc.pdf} % 99
\insertmydocument{chapter}{Conspiracy Theories and Reasonable Disagreement}{Juho Ritola}{pdfs/regular/106regularRitola.doc.pdf} % 106
\insertmydocument{chapter}{Inference to the Best Metaphor}{Cristi\'{a}n Santib\'{a}\~{n}ez}{pdfs/regular/107regularSantibanez.docx.pdf} % 107
\insertmydocument{chapter}{Beyond Randomized Clinical Trials: Emerging Innovations in Reasoning About Health}{Jodi Schneider, Sally Jackson}{pdfs/regular/109regularSchneider.docx.pdf} % 109
\insertmydocument{chapter}{Multimodal Argumentation and Dissent -- a Perspective of Multimodal Critical Discourse Analysis}{Andrea Sabine Sedlaczek}{pdfs/regular/111regularSedlaczek.doc.pdf} % 111
\insertmydocument{chapter}{Analogical Argumentation in Philosophical Thought Experiments}{Stefan Sleeuw}{pdfs/regular/112regularSleeuw.docx.pdf} % 112
\insertmydocument{chapter}{Detecting and Handling Disagreement in Multi-party Health Coaching}{Mark Snaith, Alison Pease, Dominic De Franco}{pdfs/regular/113regularSnaith.doc.pdf} % 113
\insertmydocument{chapter}{Explaining Away and the Propensity Interpretation of Probability: the Case of Unequal Priors}{Alice Liefgreen, Marko Te\v{s}i\'{c}}{pdfs/regular/116regularLiefgreen.docx.pdf} % 116
\insertmydocument{chapter}{Responsible Agent Deliberation}{Douglas Walton, Alice Toniolo}{pdfs/regular/118regularToniolo.doc.pdf} % 118
\insertmydocument{chapter}{Disagreement in Aristotle’s Topics and its Latin Medieval Reception}{Gustavo Fern\'{a}ndez Walker, Ana Mar\'{i}a Mora-M\'{a}rquez}{pdfs/regular/122regularWalker.docx.pdf} % 122
\insertmydocument{chapter}{Conductive Arguments. Towards a Solution of the Riddles}{Harald R. Wohlrapp}{pdfs/regular/125regularWohlrapp.doc.pdf} % 125
\insertmydocument{chapter}{A Ludological Perspective on the Shape of Argument: Collaborative Assent to Dissenting Opposition}{Michael A. Yong-Set}{pdfs/regular/126regularYong-Set.docx.pdf} % 126
\insertmydocument{chapter}{Towards the Justification of Warrant in a Legal Hard Case: How to Justify a Legal Rule by the Backing}{Shiyang Yu, Xi Chen}{pdfs/regular/127regularYu.docx.pdf} % 127
\insertmydocument{chapter}{How Should We Classify Argument Schemes?}{David Zarefsky}{pdfs/regular/129regularZarefsky.doc.pdf} % 129