### 2020-04-20 papers-vol-1.tex
updated titles vol 1

\part{Long papers}

\insertmydocument{chapter}{Argumentative Design and Polylogue}{Mark Aakhus}{pdfs/long/1longAakhus.docx.pdf}

\insertmydocument{chapter}{Revisiting Apologie de la Polémique: Coexistence in Dissent and its “Felicity Conditions”}{Ruth Amossy}{pdfs/keynote/1keynoteAmossy.doc.pdf}

\insertmydocument{chapter}{What is Wrong with Deductivism?}{Lilian Bermejo-Luque}{pdfs/long/3LongBermejo-Luque.doc.pdf}

\insertmydocument{chapter}{Is There an Informal Logic Approach to Argument?}{J. Anthony Blair}{pdfs/long/4longBlair.docx.pdf}
\insertmydocument{chapter}{Commentary on Blair’s Is there an informal logic approach to argument?}{Paula Olmos}{pdfs/commentary/4longBlair-C.doc.pdf}

\insertmydocument{chapter}{Structural Differences Between Practical and Cognitive Presumptions}{Petar Bodlovi\'{c}}{pdfs/long/5longBodlovic.doc.pdf}
\insertmydocument{chapter}{Commentary on Bodlovi\'{c} Structural Differences Between Practical and Cognitive Presumptions}{James B. Freeman}{pdfs/commentary/5longBodlovic-C.doc.pdf}

\insertmydocument{chapter}{Reliability of Argument Mapping}{Sebastian Cacean}{pdfs/long/6longCacean.docx.pdf}
\insertmydocument{chapter}{Commentary on Cacean’s Reliability of Argument Mapping}{Roosmaryn Pilgram}{pdfs/commentary/6longCacean-C.doc.pdf}

\insertmydocument{chapter}{Critical discussion for sub-optimal settings}{Diego Castro}{pdfs/long/7longCastro.doc.pdf}
\insertmydocument{chapter}{Comments on Diego Castro’s ``Critical Discussion in Sub-Optimal Settings''}{David Godden}{pdfs/commentary/7longCastro-C.doc.pdf}

\insertmydocument{chapter}{Democratic Legitimacy and Acts of Dissent}{Cristina Corredor}{pdfs/long/8longCorredor.docx.pdf}
\insertmydocument{chapter}{Dissent and Democracy -- Commentary on Corredor’s Democratic Legitimacy and Acts of Dissent}{David Zarefsky}{pdfs/commentary/8longCorredor-C.doc.pdf}

\insertmydocument{chapter}{Strength of Reasons for Moral Dissent}{James Freeman}{pdfs/long/9longFreeman.doc.pdf}
\insertmydocument{chapter}{What is Peculiar about Moral Dissent? Commentary on Freeman’s ``Strength of Reasons for Moral Dissent''}{John Patrick Casey}{pdfs/commentary/9longFreeman-C.docx.pdf}

\insertmydocument{chapter}{Where Are Dissent and Reasons in Epistemic Justification?}{Jos\'{e} \'{A}ngel Gasc\'{o}n}{pdfs/long/10longGascon.doc.pdf}
\insertmydocument{chapter}{Knowledge and the Epistemic Function of Argumentation -- Commentary on Gasc\'{o}n's ``Where Are dissent and Reasons in Epistemic Justification?''}{Christoph Lumer}{pdfs/commentary/10longGascon-C.doc.pdf}

\insertmydocument{chapter}{Justifying Questions?}{G.C. Goddu}{pdfs/long/11LongGoddu.doc.pdf}
\insertmydocument{chapter}{Commentary on G. C. Goddu’s ``Justifying Questions?''}{David Hitchcock}{pdfs/commentary/11longGoddu-C.docx.pdf}

\insertmydocument{chapter}{Ethnography of Argumentation}{Kati Hannken-Illjes}{pdfs/long/12longHannkenilljes.docx.pdf}

\insertmydocument{chapter}{“’I said what I said’ - Black Women and Argumentative Politeness Norms”}{Tempest M Henning}{pdfs/long/13longHenning.docx.pdf}
\insertmydocument{chapter}{Commentary on Tempest Henning’s “’I Said What I Said’ - Black Women and Argumentative Politeness Norms”}{Catherine Hundleby}{pdfs/commentary/13longHenning-C.doc.pdf}

\insertmydocument{chapter}{The Argument Assessment Tutor (AAT)}{Michael H.G. Hoffmann}{pdfs/long/14longHoffmann.doc.pdf}

\insertmydocument{chapter}{Grice, Machine Head and the Problem of Overexpressed Premises}{Michael J. Hoppmann}{pdfs/long/15longHoppmann.doc.pdf}
\insertmydocument{chapter}{Commentary on Hoppmann’s ``Grice, Machine Head and the Problem of Overexpressed Premises''}{Thierry Herman}{pdfs/commentary/15longHoppmann-C.docx.pdf}

\insertmydocument{chapter}{Standpoints and Commitments as Products of Argumentative Work: Micro/Macro- Analysis of an Infamous Press Conference}{Sally Jackson, Scott Jacobs, Xiaoqi Zhang}{pdfs/long/16longJackson.doc.pdf}
\insertmydocument{chapter}{A wanting analysis -- Commentary on S. Jackson, S. Jacobs, Y. Zhang's ``Standpoints and Commitments as Products of Argumentative Work: Micro/Macro-Analysis of an Infamous Press Conference''}{Harald R. Wohlrapp}{pdfs/commentary/16longJackson-C.doc.pdf}

\insertmydocument{chapter}{Speech Act Pluralism in Argumentative Polylogues}{Marcin Lewi\'{n}ski}{pdfs/long/18longLewinski.doc.pdf}

\insertmydocument{chapter}{Evaluating Relevance in Analogical Arguments Through Warrant-based Reasoning}{John Licato, Michael Cooper}{pdfs/long/19longLicato.doc.pdf}
\insertmydocument{chapter}{Commentary on Licato {\&} Cooper's Evaluating Relevance in Analogical Arguments Through Warrant-based Reasoning}{Nicole Cruz}{pdfs/commentary/19longLicato-C.docx.pdf}

\insertmydocument{chapter}{Arguments From Expert Opinion – An Epistemological Approach}{Christoph Lumer}{pdfs/long/20longLumer.doc.pdf}

\insertmydocument{chapter}{All-out Attack}{Tshepo Bogosi Mosaka}{pdfs/long/21longMosaka.doc.pdf}

\insertmydocument{chapter}{Strategic Maneuvering with Speech Codes: The Rhetorical Use of Cultural Presumptions in Constructing Argumentative Discourse}{Menno H. Reijven}{pdfs/long/22longReijven.doc.pdf}
\insertmydocument{chapter}{Commentary on Reijven’s Strategic Maneuvering with Speech Codes: The Rhetorical Use of Cultural Presumptions in Constructing Argumentative Discourse}{Beth Innocenti}{pdfs/commentary/22longReijven-C.doc.pdf}

\insertmydocument{chapter}{Disagreement, Public Reasoning, and (Non-)Authoritarian Argumentation}{Karim Sadek}{pdfs/long/23longSadek.doc.pdf}
\insertmydocument{chapter}{Commentary on Sadek’s Disagreement, Public Reasoning, and (Non-)Authoritarian Argumentation}{Cristina Corredor}{pdfs/commentary/23longSadek-C.doc.pdf}

\insertmydocument{chapter}{Uncovering Hidden Premises to Reveal the Arguer's Implicit Values: Analysing the Public Debate About Funding Prep}{Sean Sinclair}{pdfs/long/24longSinclair.doc.pdf}
\insertmydocument{chapter}{Commentary on Sinclair’s Uncovering Hidden Premises to Reveal the Arguer's Implicit Values: Analysing the Public Debate about Funding Prep}{Sally Jackson}{pdfs/commentary/24longSinclair-C.doc.pdf}

\insertmydocument{chapter}{Strange Fish: Belief and the Roots of Disagreement}{Christopher W. Tindale}{pdfs/long/25longTindale.docx.pdf}
\insertmydocument{chapter}{An optimistic perspective on deep disagreement? Commentary on Tindale’s Strange Fish: Belief and the Roots of Disagreement}{L\'{e}a Farine}{pdfs/commentary/25longTindale-C.doc.pdf}

\insertmydocument{chapter}{Profiling Dialogues: Multi-trait Mapping of Televised Argumentative Exchanges}{G\'{a}bor \'{A} Zempl\'{e}n}{pdfs/long/27longZemplen.doc.pdf}
\insertmydocument{chapter}{Commentary on Zempl\'{e}n’s Profiling Dialogues: Multi-trait Mapping of Televised Argumentative Exchanges}{Mehmet Ali \"{U}zelg\"{u}n}{pdfs/commentary/27longZemplen-C.doc.pdf}

\part{Thematic papers}

\insertmydocument{chapter}{Visual Argument Schemes in the PTA}{Leo Groarke}{pdfs/thematic/1regularthematicGroarke.docx.pdf}
\insertmydocument{chapter}{The Prospects for Multimodal Schemes of Argument: Assessing the Spoofing Strategies in Subvertisements of the Tobacco Industry
}{Assimakis Tseronis}{pdfs/thematic/3regularThematicTseronis.docx.pdf}
\insertmydocument{chapter}{(Mis)Framing Photographs as an Obstacle to Fair Dissent}{Marta Zampa, Chiara Pollaroli}{pdfs/thematic/4regularThematicZampa.docx.pdf}
\insertmydocument{chapter}{Dissent, Disadvantage, Testimony and the Ideological ‘Truth’ of Presumptions}{Harmony Peach}{pdfs/thematic/5regularThematicPeach.doc.pdf}
\insertmydocument{chapter}{With All Due Respect: Controversial Beliefs and the Limits of Responsible Argumentation}{Tracy Bowell}{pdfs/thematic/7regularThematicBowell.doc.pdf}
\insertmydocument{chapter}{A Quantitative Corpus-Based Study of Evidentiality and Disagreement in Earnings Conference Calls}{Andrea Rocci, Carlo Raimondo, Daniele Puccinelli}{pdfs/thematic/9regularThematicRocci.docx.pdf}
\insertmydocument{chapter}{Five Approaches to Argument Strength: Probabilistic, Dialectical, Structural, Empirical, and Computational}{Frank Zenker, David Godden, Marcin Selinger, Simon Wells, Kamila Debowska-Kozlowska}{pdfs/thematic/13regularThematicZenkeretal.docx.pdf}
\insertmydocument{chapter}{Responsible Agent Deliberation}{Douglas Walton, Alice Toniolo}{pdfs/thematic/18regularThematicWalton.doc.pdf}
\insertmydocument{chapter}{From Dialogic to Argument-Based Teaching: Introducing Pragmatic Criteria to Analyse Whole-Class Interactions}{Chrysi Rapanta, Dilar Cascalheira}{pdfs/thematic/20regularThematicRapanta.doc.pdf}