### 2020-05-10 prep.tex

\newpage

\section*{Conference Organization}
	\addcontentsline{toc}{chapter}{Conference Organization}
	\subsection*{Organising committee}
	\begin{compactitem}[]
		\item[] Catarina Dutilh Novaes (Vrije Universiteit Amsterdam)
		\item[] Henrike Jansen (University of Leiden)
		\item[] Jan Albert Van Laar (chair) (University of Groningen)
		\item[] Bart Verheij (University of Groningen)
	\end{compactitem}

	\subsection*{Programme committee}
	\begin{compactitem}[]
		\item[] Dale Hample (chair) (University of Maryland)
		\item[] Henrike Jansen (secretary) (University of Leiden)
		\item[] Marcin Lewinski (Universidade Nova de Lisboa)
		\item[] Juho Ritola (University of Turku)
		\item[] Frank Zenker (Lund University)
 	\end{compactitem}

 	\subsection*{ECA steering committee}
 	\begin{compactitem}[]
		\item[] Lilian Bermejo Luque (University of Granada)
		\item[] Katarzyna Budzyńska (Polish Academy of Sciences)
		\item[] Henrike Jansen (University of Leiden)
		\item[] Marcin Koszowy (University of Białystok)
		\item[] Jan Albert Van Laar (deputy chair) (University of Groningen)
		\item[] Marcin Lewiński (Universidade Nova de Lisboa)
		\item[] Dima Mohammed (Universidade Nova de Lisboa)
		\item[] Steve Oswald (University of Fribourg)
		\item[] Fabio Paglieri (chair) (ISTC-CNR)
		\item[] Juho Ritola (University of Turku)
		\item[] Sara Rubinelli (University of Lucerne)
		\item[] Frank Zenker (University of Lund)
	\end{compactitem}

	\subsection*{Scientific Panel}
	\begin{compactitem}[]
		\item[] Mark Aakhus (Rutgers University)
		\item[] Andrew Aberdein (Florida Institute of Technology)
		\item[] Scott Aikin (Vanderbilt University)
		\item[] Mehmet Ali Uzelgun (Universidade NOVA de Lisboa)
		\item[] Ruth Amossy (Tel-Aviv University)
		\item[] Corina Andone (University of Amsterdam)
		\item[] Michael Baumtrog (Ryerson University)
		\item[] Gregor Betz (KIT Karlsruher)
		\item[] Sarah Bigi (Università Cattolica del Sacro Cuore)
		\item[] Anthony Blair (University of Windsor)
		\item[] Patrick Bondy (Cornell University)
		\item[] Alessandra Von Burg (Wake Forest University)
		\item[] John Casey (Northeastern Illinois University)
		\item[] Daniel Cohen (Colby College)
		\item[] Louis De Saussure (University of Neuchatel)
		\item[] Marianne Doury (Paris Descartes University)
		\item[] Michel Dufour (Université Sorbonne-Nouvelle)
		\item[] Isabela Fairclough (University of Cantral Lancashire)
		\item[] Eveline Feteris (University of Amsterdam)
		\item[] Joana Garmendia (University of the Basque Country)
		\item[] Bart Garssen (University of Amsterdam)
		\item[] Ingeborg van der Geest (University of Amsterdam)
		\item[] David Godden (Michigan State University)
		\item[] C. Goddu (University of Richmond)
		\item[] Jean Goodwin (North Carolina State University)
		\item[] Floriana Grasso (University of Liverpool)
		\item[] Sara Greco (Università della Svizzera Italiana)
		\item[] Leo Groarke (Trent University)
		\item[] Thierry Herman (Universities of Neuchatel and lausanne)
		\item[] Martin Hinton (University of Lodz)
		\item[] David Hitchcock (McMaster University)
		\item[] Hans Hoeken (Utrecht University)
		\item[] Michael Hoffmann (Georgia Institute of Technology)
		\item[] Michael Hoppmann (Northeastern University)
		\item[] Jos Hornikx (Radboud University)
		\item[] Catherine Hundleby (University of Windsor)
		\item[] Constanza Ihnen (University of Chile, Law Faculty, Institute of Argumentation)
		\item[] Beth Innocenti (University of Kansas)
		\item[] Jérôme Jacquin (University of Lausanne)
		\item[] Henrike Jansen (University of Leiden)
		\item[] Amy Johnson (University of Oklahoma)
		\item[] Manfred Kienpointner (University of Innsbruck)
		\item[] Gabrijela Kisicek (University of Zagreb)
		\item[] Susan Kline (The Ohio State University)
		\item[] Bart van Klink (Vrije Universiteit Amsterdam)
		\item[] Manfred Kraus (University of Tübingen)
		\item[] Maarten van Leeuwen (Leiden University)
		\item[] Christoph Lumer (University of Siena)
		\item[] Fabrizio Macagno (Universidade NOVA de Lisboa)
		\item[] Andrei Moldovan (University of Salamanca)
		\item[] Paula Olmos (Universidad Autónoma de Madrid)
		\item[] Rudi Palmieri (University of Liverpool)
		\item[] Steven Patterson (Marygrove College)
		\item[] Niki Pfeifer (Munich Center for Mathematical Philosophy)
		\item[] Gilbert Plumer (Law School Admission Council) (retired)
		\item[] Lotte van Poppel (University of Amsterdam)
		\item[] Henry Prakken (University of Groningen)
		\item[] Martin Reisigl (University of Bern)
		\item[] Adam Richards (Texas Christian University)
		\item[] Juho Ritola (University of Helsinki)
		\item[] Andrea Rocci (Università della Svizzera Italiana)
		\item[] Harvey Siegel (University of Miami)
		\item[] Francisca Snoeck Henkemans (University of Amsterdam)
		\item[] Manfred Stede (University of Potsdam)
		\item[] Katharina Stevens (University of Lethbridge)
		\item[] Christopher Tindale (University of Windsor)
		\item[] Assimakis Tseronis (Örebro University)
		\item[] Mariusz Urbanski (Adam Mickiewicz University)
		\item[] Hans Vilhelm Hansen (University of Windsor)
		\item[] Laura Vincze (Universityt of Pisa)
		\item[] Jacky Visser (University of Dundee)
		\item[] Lena Wahlberg (Lund University)
		\item[] Harry Weger (University of Central Florida)
		\item[] David Williams (Florida Atlantic University)
		\item[] Harald Wohlrapp (Universität Hamburg)
		\item[] Carly Woods (University of Maryland)
		\item[] Yun Xie (Sun Yat-sen University)
		\item[] Minghui Xiong (Sun Yat-sen University)
		\item[] Igor Z. Zagar (Educational Research Institute)
		\item[] Marta Zampa (Zurich University of Applied Sciences)
		\item[] David Zarefsky (Northwestern University)
		\item[] Gabor Zemplen (Budapest University of Technology and Economics)
	\end{compactitem}