### 2020-05-12 papers-vol-3.tex

\part{Regular papers}

\insertmydocument{chapter}{The Spaces and Places of Argumentation: Developing a Spatial Model of Argumentation}{Steven Johnson}{pdfs/regular/58regularJohnson.docx.pdf} % 58
\insertmydocument{chapter}{Aggressiveness as a Qualifier of Dissent in Interpersonal Arguing}{Iryna Khomenko}{pdfs/regular/59regularKhomenko.docx.pdf} % 59
\insertmydocument{chapter}{Auditory Arguments – Importance of Sound in an Argumentative Discourse
(An Empirical Study)}{Gabrijela Ki\v{s}i\v{c}ek}{pdfs/regular/60regularKisicek.docx.pdf} % 60
\insertmydocument{chapter}{Be reasonable! Ways to React to Cases of Presumed Unreason}{Erik C. W. Krabbe, Jan Albert van Laar}{pdfs/regular/65regularKrabbe.doc.pdf} % 65
\insertmydocument{chapter}{Epistemic Injustice and Deep Disagreement}{Thirza Lagewaard}{pdfs/regular/69regularLagewaard.doc.pdf} % 69
\addtocontents{toc}{\protect\newpage}
\insertmydocument{chapter}{Argumentative Patterns of Right-Wing Populism: The Example of the “Alternative for Germany”}{David Lanius}{pdfs/regular/70regularLanius.doc.pdf} % 70
\insertmydocument{chapter}{A Decision Tree for Annotating Argumentation Scheme Corpora}{John Lawrence, Jacky Visser, Douglas Walton, Chris Reed}{pdfs/regular/71regularLawrence.doc.pdf} % 71
\insertmydocument{chapter}{Philosophical Foundation of Reasonableness in Mencius’s Argumentative Discourse: Based on the use of dissociation}{Yan Linqiong, Xiong Minghui}{pdfs/regular/72regularYanXiong.docx.pdf} % 72
\insertmydocument{chapter}{Strategic Manoeuvring with the Construction “We (All) Know that X” in Parliamentary Debates}{Matthijs Looij, Bertijn van der Steenhoven}{pdfs/regular/74regularLooij.docx.pdf} % 74
\insertmydocument{chapter}{Argument, Justification and Inquiry: a Pragmatist Approach}{Federico E. L\'{o}pez}{pdfs/regular/75regularLopez.doc.pdf} % 75
\insertmydocument{chapter}{Dissent as a Voice of Support: The Influence of Argument Types in Providing Support to Those Who have Experienced Racial Discrimination}{Uttara Manohar, Susan L. Kline}{pdfs/regular/78regularManohar.doc.pdf} % 78
\insertmydocument{chapter}{Holism of Reasons and its Consequences for Argumentation Theory}{Hubert Marraud}{pdfs/regular/80regularMarraud.doc.pdf} % 80
\insertmydocument{chapter}{“Why did not your correspondent make an honest inquiry before so writing?”: The Text Structure and Discourse of Disagreement in Irish letters to the Editor}{Davide Mazzi}{pdfs/regular/81regularMazzi.doc.pdf} % 81
\insertmydocument{chapter}{An Argumentative Approach to the Justification of Abduction}{Paula Olmos}{pdfs/regular/86regularOlmos.doc.pdf} % 86
\insertmydocument{chapter}{When You Can’t Trust What You Don’t Know: Organ and Tissue Donation, Public Trust, the Voice of Dissent and Informed Consent}{Harmony Peach}{pdfs/regular/91regularPeach.doc.pdf} % 91
\insertmydocument{chapter}{“The worst ever conceived by a man of genius. ” Hume’s Probability Argument in A Treatise}{Jeanne Peijnenburg, David Atkinson}{pdfs/regular/92regularPeijnenburg.docx.pdf} % 92
\insertmydocument{chapter}{Character, Dog Whistles, and the Limits of Charity}{Kathryn Phillips}{pdfs/regular/94regularPhillips.doc.pdf} % 94
\insertmydocument{chapter}{The Rhetorical Potential of Metaphor: An Experimental Study on the Effect of Metaphors in Argumentation}{Roosmaryn Pilgram}{pdfs/regular/97regularPilgram.docx.pdf} % 97
\insertmydocument{chapter}{Separate Opinions as an Argumentative Activity Type}{H. Jos\'{e} Plug}{pdfs/regular/98regularPlug.doc.pdf} % 98
\insertmydocument{chapter}{Can Literary Fiction be Suppositional Reasoning?}{Gilbert Plumer}{pdfs/regular/99regularPlumer.doc.pdf} % 99
\insertmydocument{chapter}{Conspiracy Theories and Reasonable Disagreement}{Juho Ritola}{pdfs/regular/106regularRitola.doc.pdf} % 106
\insertmydocument{chapter}{Inference to the Best Metaphor}{Cristi\'{a}n Santib\'{a}\~{n}ez}{pdfs/regular/107regularSantibanez.docx.pdf} % 107
\insertmydocument{chapter}{Beyond Randomized Clinical Trials: Emerging Innovations in Reasoning About Health}{Jodi Schneider, Sally Jackson}{pdfs/regular/109regularSchneider.docx.pdf} % 109
\insertmydocument{chapter}{Multimodal Argumentation and Dissent -- a Perspective of Multimodal Critical Discourse Analysis}{Andrea Sabine Sedlaczek}{pdfs/regular/111regularSedlaczek.doc.pdf} % 111
\insertmydocument{chapter}{Analogical Argumentation in Philosophical Thought Experiments}{Stefan Sleeuw}{pdfs/regular/112regularSleeuw.docx.pdf} % 112
\insertmydocument{chapter}{Detecting and Handling Disagreement in Multi-party Health Coaching}{Mark Snaith, Alison Pease, Dominic De Franco}{pdfs/regular/113regularSnaith.doc.pdf} % 113
\insertmydocument{chapter}{Explaining Away and the Propensity Interpretation of Probability: the Case of Unequal Priors}{Alice Liefgreen, Marko Te\v{s}i\'{c}}{pdfs/regular/116regularLiefgreen.docx.pdf} % 116
\insertmydocument{chapter}{Responsible Agent Deliberation}{Douglas Walton, Alice Toniolo}{pdfs/regular/118regularToniolo.doc.pdf} % 118
\insertmydocument{chapter}{Disagreement in Aristotle’s Topics and its Latin Medieval Reception}{Gustavo Fern\'{a}ndez Walker, Ana Mar\'{i}a Mora-M\'{a}rquez}{pdfs/regular/122regularWalker.docx.pdf} % 122
\insertmydocument{chapter}{Conductive Arguments. Towards a Solution of the Riddles}{Harald R. Wohlrapp}{pdfs/regular/125regularWohlrapp.doc.pdf} % 125
\insertmydocument{chapter}{A Ludological Perspective on the Shape of Argument: Collaborative Assent to Dissenting Opposition}{Michael A. Yong-Set}{pdfs/regular/126regularYong-Set.docx.pdf} % 126
\insertmydocument{chapter}{Towards the Justification of Warrant in a Legal Hard Case: How to Justify a Legal Rule by the Backing}{Shiyang Yu, Xi Chen}{pdfs/regular/127regularYu.docx.pdf} % 127
\insertmydocument{chapter}{How Should We Classify Argument Schemes?}{David Zarefsky}{pdfs/regular/129regularZarefsky.doc.pdf} % 129